Boolean Operations for Surfaces in VTK Without External Libraries

### 2011-05-11 Document/LaTeX/Article.tex
Reworked section on intersection line splitting.

% The abstract should be a paragraph or two long, and describe the
% scope of the document.
\begin{abstract}
\noindent
We have written a set of classes than enable boolean operations on the space contained within surface meshes defined in \code{vtkPolyData} objects using only VTK classes. In addition to being compatible with the VTK license, our contribution preserves mesh topology and passes point data and cell data through to outputs.

@@

\begin{figure}
\centering
\includegraphics[scale=0.25]{Figures/BoxNormals}
\itkcaption{Splitting a mesh with duplicate normals. The box object has three duplicate points at each corner each of which has a different normal (gray arrows). The surface intersection between the box in the front (gray, solid) and the box in the back (white, outline) is shown in green. Points from the intersection lines that lie on the duplicate edges of the box are also duplicated by the mesh splitting procedure in \code{vtkPolyDataIntersection}. The normal at a duplicated point, along with other point data, is interpolated from the cell that contains the intersection line that references the point.}
\label{fig:BoxNormals}
\end{figure}

XXXX work on below paragraph again

More generally, the end result of splitting the surface intersection should be that a copy of each line end point is produced for each set of triangles that share an edge. This can be achieved by iterating over the key-value pairs in the multimap, appending a copy of the point corresponding to the point ID in the key to the end of the \code{vtkPoints} object storing the endpoints, and removing all records in the multimap where the point ID is the same as the first point and the cell ID marks a cell that shares the edge identified in the first record. The point indices of the lines within the cells that share an edge are renumbered to the ID of the new point in the \code{vtkPoints} object. Finally, the point ID and the cell ID from one of the cells that shares the edge is stored in a point-to-cell map for future use in interpolating point data at the new point. It does not matter which cell ID is associated with a point on the edge of a cell because the interpolation results will be the same.

The points in the input mesh and their associated point data are copied to the points in the output mesh. The points from the intersection lines are then appended to the point set in the output mesh, and the point data is interpolated at those points.

@@

\subsection{Setting up a VTK Pipeline}

@@

\code{vtkThreshold} \code{vtkThreshold}

$\downarrow$ \hspace{5em}  $\downarrow$

\code{vtkDataSetSurfaceFilter} \code{vtkDataSetSurfaceFilter}

@@

When computing the difference boolean operation, part of the resulting surface mesh will have reversed normals and a reversed order of cell points. The ReorientDifferenceCells option is used to enable reversal of the normals and reorientation of these cells and can be set using the methods
\begin{verbatim}
void SetReorientDifferenceCells(int value);
void ReorientDifferenceCellsOn();
void ReorientDifferenceCellsOff();
\end{verbatim}

Lastly, the Tolerance sets a threshold on the absolute value of a distance field value below which the distance is considered to be zero.
\section{Results}

>>>>

% The abstract should be a paragraph or two long, and describe the
% scope of the document.
\begin{abstract}
\noindent
We have written a set of classes than enable computation of boolean operations on surface meshes using only VTK classes. In addition to being compatible with the VTK license, our contribution preserves surface mesh topology (to the extent possible) and passes point data and cell data through to the output mesh where possible.

@@

\begin{figure}
\centering
\includegraphics[scale=0.25]{Figures/BoxNormals}
\itkcaption{Splitting a mesh with duplicate normals. The box object has three duplicate points at each corner each of which has a different normal (gray arrows). The surface intersection between the box in the front (gray, solid) and the box in the back (white, outline) is shown in green. Points from the intersection lines that lie on the duplicate edges of the box are also duplicated by the mesh splitting procedure in \code{vtkPolyDataIntersection}. The normal at a duplicated point, along with other point data, is interpolated from the cell that contains the intersection line that references that point.}
\label{fig:BoxNormals}
\end{figure}

More generally, the end result of splitting the surface intersection should be that a copy of each line end point is produced for each set of triangles that share an edge. This is done by iterating over the entries of the edge point multimap. The first time a particular point ID is encountered in a multimap key-value pair, all other entries with the same key are checked to determine if the cell with the cell ID in the value shares the edge indicated by the first entry's value. If the entry is for a shared edge, then the entry is removed from the multimap and not processed any further. If not, the entry is left in the multimap for further processing.

If any multimap entries are left that have the current point ID as key, then that point lies on a split in the mesh and requires duplication. In this case, a copy of the point is inserted into the list of points for the intersection line data structure, and the first such entry in the multimap is read. All other entries in the multimap with the point ID as key are again examined to determine which share an edge with the first entry. Entries that do not share an edge are left in the multimap. For entries that share an edge, the point ID in the line referenced by the line ID stored in the entry is replaced with the ID of the newly duplicated point. The entry is then removed from the multimap. The process described in this paragraph is repeated if any entries are left that have the current point ID as key. Otherwise, the first multimap entry is removed.

During the above process, another auxiliary data structure is built. Every time a point is duplicated, a key-value pair is added to a map where the key is the point ID and the value is the cell ID from which the point data at the point should be interpolated. It does not matter which cell ID from the set of cells that share an edge is associated with a point on that edge because the interpolation results will be the same.

After the intersection lines are split, the points in the input mesh and their associated point data are copied to the points in the output mesh. The points from the intersection lines are then appended to the point set in the output mesh, and the point data is interpolated at those points.

@@

\subsection{Setting up a VTK Pipeline for Boolean Operations on Surfaces}

@@

\code{vtkThreshold} \hspace{1em} \code{vtkThreshold}

$\downarrow$ \hspace{6em}  $\downarrow$

\code{vtkDataSetSurfaceFilter} \hspace{1em} \code{vtkDataSetSurfaceFilter}

@@

When computing the difference boolean operation, part of the resulting surface mesh will have reversed normals and a reversed order of cell points. The \code{ReorientDifferenceCells} option is used to enable reversal of the normals and reorientation of these cells and can be set using the methods
\begin{verbatim}
void SetReorientDifferenceCells(int value);
void ReorientDifferenceCellsOn();
void ReorientDifferenceCellsOff();
\end{verbatim}

Lastly, the \code{Tolerance} sets a threshold on the absolute value of a distance field value below which the distance is considered to be zero.
\section{Results}
